---
title: 'Crop-Yield analysis: Dashboard'
author: ''
date: '2020-01-06'
slug: crop-yield-analysis-dashboard
categories: []
tags:
  - R Shiny Apps
  - Statistics
subtitle: ''
summary: ''
authors: []
lastmod: '2021-03-28T17:29:55+03:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
runtime: shiny
output:
  blogdown::html_page:
    toc: false
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
```

In this blog we look at the crop-yields in different countries from 1961 to 2018. Crop Yields is in tonnes per hectare

The crops include bananas, barley, beans, cassava, cocoa beans, maize, peas, potatoes, rice, soya beans and wheat.

We first begin by loading the packages and then setting the theme.

```{r eval=TRUE, include=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(shiny)
library(stringr)
library(plotly)

theme_set(theme_light())
```

## Data cleaning and exploration

We find out what the top 9 crops in the data sets.



```{r echo=TRUE}
# Data cleaning and processing
yields_tidy <- readRDS("yields_tidy.rds")
top_nine <- yields_tidy %>%
  count(crop, sort = TRUE) %>%
  head(9) 

top_crops <- top_nine%>%
  pull(crop)

top_nine %>% knitr::kable()
```

We have a look at the crop yield of Kenya and Tanzania. We can see that Cassava, Beans and Potatoes are the most produced crops.

```{r echo=TRUE}
yields_tidy %>% 
  filter(code %in% c("TZA", "KEN")) %>%
    mutate(crop = fct_reorder(crop, -yield)) %>%
    mutate(entity = fct_reorder(entity, -yield)) %>%
    ggplot(aes(year, yield, color = entity)) +
    geom_line() +
    expand_limits(y = 0) +
    facet_wrap(~ crop, scales = "fixed") +
    labs(x = "Year",
         y = "Yield (tonnes per hectare)",
         title = "Crop yields over time",
         color = "Country")
```


```{r}
plot_yields <- function(tbl, facet_scales = "fixed") {
  g <- tbl %>%
    mutate(crop = fct_reorder(crop, -yield)) %>%
    mutate(entity = fct_reorder(entity, -yield)) %>%
    ggplot(aes(year, yield, color = entity)) +
    geom_line() +
    expand_limits(y = 0) +
    facet_wrap(~ crop, scales = facet_scales) +
    labs(x = "Year",
         y = "Yield (tonnes per hectare)",
         title = "Crop yields over time",
         color = "Country")
  
  ggplotly(g)
}
# yields_tidy %>%
#   filter(code %in% c("RWA", "SEN")) %>%
#   plot_yields()
```


Here is an example of the result with one of my Shiny app ([link](https://josabusmart.shinyapps.io/World_crops/) to the app):

<iframe height="800" width="100%" frameborder="no" src="https://josabusmart.shinyapps.io/World_crops/"> </iframe>

<br>

The code to the dashboard is shown below.

```{r eruptions, eval=FALSE, include=TRUE}

ui <- fluidPage(
  
  titlePanel("Crop Yields of all Countries/Continents/Region from 1961 to 2018: Victor Mandela"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("entity",
              label = "Country Name:",
              choices = unique(yields_tidy$entity),
              selected = c("Kenya", "Tanzania"),
              selectize = TRUE,
              multiple = TRUE),
      selectInput("crop",
              label = "Types of Crops:",
              choices = unique(yields_tidy$crop),
              selected = top_crops,
              selectize = TRUE,
              multiple = TRUE),
      radioButtons("facet_scales",
               label = "",
               choices = c("Free y-axis" = "free_y",
                           "Shared y-axis" = "fixed"))
      ),
      mainPanel(
  plotlyOutput("crop_plot"),
  dataTableOutput("tabled")
      )
    )
  )


server <- function(input, output){
  plot_yields <- function(tbl, facet_scales = "fixed") {
  g <- tbl %>%
    mutate(crop = fct_reorder(crop, -yield)) %>%
    mutate(entity = fct_reorder(entity, -yield)) %>%
    ggplot(aes(year, yield, color = entity)) +
    geom_line() +
    expand_limits(y = 0) +
    facet_wrap(~ crop, scales = facet_scales) +
    labs(x = "Year",
         y = "Yield (tonnes per hectare)",
         title = "Crop yields over time",
         color = "Country",
         caption = "Data source: TidyTuesday")
  
  ggplotly(g)
  }
  
output$crop_plot <- renderPlotly({
  yields_tidy %>%
    filter(entity %in% input$entity,
           crop %in% input$crop) %>%
    plot_yields(facet_scales = input$facet_scales)
})

output$tabled <- renderDataTable(yields_tidy,
                                 options = list(
          pageLength = 5
        )
      )

}
shinyApp(ui = ui, server = server)
```
